---
title: "Wastewater, Rate of change of positive tests"
author: "`Marlin derived from work by Brian Yandell"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 4,
	fig.width = 15,
	message = FALSE,
	warning = FALSE
)
```

```
The code was derived from work by the DSI. This analysis looks at the rate of change of the positive rate of COVID-19 in Madison Wisconsin and matches this with the wastewater data. While offering promising results Difficulties arise when trying to convert it to a measure of positive COVID-19 rates. This and bad fundamentals leads to the conclusion that this is a bad model for Covid-19 positive rates.
The Original R code file can be found in the [pandemic github repository](https://github.com/UW-Madison-DataScience/pandemic/blob/master/wastewater.Rmd).
The Code for this R File can be found in the [Covid Waste Water Exploration](https://github.com/MarlinRLee/Covid-Waste-Water-Exploration/blob/main/general%20model%20finding.Rmd)
```




```{r}
library(tidyverse)
library(readxl)
library(broom)
library(zoo)
library(RcppRoll)
library(ggformula)
library(ggpubr)
kprint <- function(...) {
  if(interactive()) {
    print(...)
  } else {
    knitr::kable(...)
  }
}
```

```{r 2}
filename <- "../data/wastewater/WW SARS-COV-2 Data V3.xlsx"
#filename <- "data/wastewater/UW and MMSD Report for SARS-Cov-2 Influent Samples 12.08.20vFinal.xlsx"
```

```{r warning = FALSE, message = FALSE}
tmpfn <- function(N1, N2) {
  AVG <- sqrt(N1 * N2)
  AVG <- ifelse(is.na(N1), N2, AVG)
  AVG <- ifelse(is.na(N2), N1, AVG)
}
missing_codes <- c("","NA","0","Undetected","Not Detected",
                   "Field Parameters to be filled in", 
                   "Inhibited-to be re-ran", "#DIV/0!")
sheets <- excel_sheets(filename)
water_MMSD <- read_excel(filename,
                    na = missing_codes,
                    col_types = c("text", "date", rep("numeric", 9), "text"),
                    sheet = 1) %>%
  rename(Comment = "...12")
water_Interceptors <- read_excel(filename,
                    na = missing_codes,
                    col_types = c("text", "date", rep("numeric", 8), "text"),
                    sheet = 2) %>%
  rename(Comment = "...11") %>%
  mutate(TSS = NA,
         Site = ifelse(Site == "MMSD P02", "MMSD P2", Site),
         Site = ifelse(Site == "MMSD P07", "MMSD P7", Site),
         Site = ifelse(Site == "MMSD P08", "MMSD P8", Site)) %>%
  select(1:5, TSS, everything())
water_UW_Dorms <- read_excel(filename,
                    na = missing_codes,
                    col_types = c("text", "date", rep("numeric", 9), rep("text", 1)),
                    sheet = 3) %>%
  rename(Comment = "...12")
water <- 
  bind_rows(water_MMSD,
            water_Interceptors,
            water_UW_Dorms) %>%
  rename(Date = "Collection Date",
         pH = "pH (SU)",
         Total_Flow = "Total Flow (MGD)",
         Conductivity = "Conductivity (uS/CM@25C)",
         N1 = "N1 (GC/L)",
         N2 = "N2 (GC/L)",
         PMMoV = "PMMoV (GC/L)",
         Pct_BCoV = "% Recovery (BCoV)") %>%
  filter(!is.na(Site)) %>%
  mutate(AVG = tmpfn(N1, N2),
         wt = 2 - is.na(N1) - is.na(N2),
         Site = ifelse(Site == "UW-D", "UW-LakeShore", Site),
         Site = ifelse(Site == "UW-S", "UW-Sellery", Site))
```

```{r}
waters <- water %>%
  pivot_longer(Conductivity:Pct_BCoV, names_to = "assay", values_to = "values") %>%
  filter(!is.na(values)) %>%
  mutate(wt = ifelse(assay == "AVG", wt, 1))
```

```{r}
predfn <- function(x) {
  x$pred <- 10^predict(lm(log10(values) ~ Date + Site, x, weight = wt))
  x
}
addpred <- function(x) {
  bind_rows(
    map(
      split(x, x$assay),
      predfn),
    .id = "assay")
}
```

```{r}
fitfn <- function(x, add = FALSE, ...) {
  if(add) {
    form <- formula(log10(values) ~ Date + Site)
    elems <- 2:3
    test <- "drop1"
  } else {
    form <- formula(log10(values) ~ Date * Site)
    elems <- 3
    test <- "anova"
  }
  fit <- lm(form, x, weight = wt)
  switch(test,
         anova = broom::tidy(anova(fit)[elems,]),
         drop1 = broom::tidy(drop1(fit, fit, test = "F")[elems,])) %>%
    select(term, statistic, p.value) %>%
    mutate(statistic = round(statistic, 2),
           p.value = signif(p.value, 3))
}
sumfn <- function(x, ...) {
  bind_rows(
    map(
      split(x, x$assay),
      fitfn, ...),
    .id = "assay") %>%
    arrange(term, assay)
}
timefit <- function(x) {
  bind_rows(
    sumfn(x, TRUE),
    sumfn(x)) %>%
  select(-statistic) %>%
  pivot_wider(names_from = "term", values_from = "p.value") %>%
  kprint()
}
```



```{r}
ShiftedAmount=5

print(paste("Time shifted",ShiftedAmount,"Days"))

CovidFileName = "covid.csv"
covidData = read.csv(CovidFileName)%>%
  mutate(ServiceID = ifelse(ServiceID=="MMSD","Madison",paste("MMSD P",ServiceID,sep="")),Date = as.Date(Date , format = "%m/%d/%y"))%>%
  rename(Site=ServiceID)
water2=water%>%
  mutate(Date=as.Date(Date))%>%
  mutate(Date=Date+ShiftedAmount)%>%
  mutate(Model=log(N1/(Pct_BCoV*PMMoV)))

print("Model is log(N1/(Pct_BCoV*PMMoV))")
         
MergedData=left_join(covidData,water2,by = c("Date","Site"))%>%
  filter(Site!="UW-LakeShore",Site!="UW-Sellery")


StartTesting=as.Date("2020-4-11")
EndTesting=as.Date("2021-1-4")
StartFeces=as.Date("2020-9-20")+ShiftedAmount
EndFeces=as.Date("2020-11-25")+ShiftedAmount



AdjustedData=MergedData%>%
  group_by(Site)%>%
  mutate(roll=rollmean(Cases/Tests,k = 7, fill = NA))%>%
  mutate(der=c(diff(roll),NA))


minSew=min(AdjustedData$Model,na.rm=TRUE)
maxSew=max(AdjustedData$Model,na.rm=T)
minTest=min(AdjustedData$der,na.rm=T)
maxTest=max(AdjustedData$der,na.rm=T)
sites=unique(AdjustedData$Site)


for (i in 1:length(sites)){
  tot =AdjustedData%>%
    filter(Site==sites[i])%>%
    ggplot()+aes(x=Date)+
    geom_point(aes(y=roll))+
    geom_vline(xintercept = StartFeces)+
    geom_vline(xintercept = EndFeces)+
    ylab("% of Covid Tests Positive")+
    scale_x_date(limits = c(StartTesting,EndTesting))
    ggtitle(sites[i])
  dir =AdjustedData%>%
    filter(Site==sites[i])%>%
    ggplot()+aes(x=Date)+
    geom_point(aes(y=der))+
    geom_vline(xintercept = StartFeces)+
    geom_vline(xintercept = EndFeces)+
    ylab("Diffrence of positive rate from last week")+
    scale_x_date(limits = c(StartTesting,EndTesting))+
    ylim(minTest,maxTest)
  zoom = AdjustedData%>%
    filter(Site==sites[i])%>%
    filter(!is.na(Cases/Tests))%>%
    ggplot()+aes(x=Date,y=der)+
    geom_point()+
    geom_smooth(se=FALSE)+
    geom_vline(xintercept = StartFeces)+
    geom_vline(xintercept = EndFeces)+
    ylab("Diffrence of positive rate from last week")+
    scale_x_date(limits = c(StartFeces,EndFeces))+
    ylim(minTest,maxTest)
  mod=AdjustedData%>%
    filter(Site==sites[i])%>%
    ggplot()+aes(x=Date,y=Model)+
    geom_smooth(se=FALSE)+
    geom_point()+
    geom_vline(xintercept = StartFeces)+
    geom_vline(xintercept = EndFeces)+
    ylab("log(N1 gene/(Pct_BCoV*PMMoV))")+
    scale_x_date(limits = c(StartFeces,EndFeces))+
    ylim(minSew,maxSew)
    fig=ggarrange(tot,dir, zoom, mod,nrow = 1)
    print(annotate_figure(fig,top = text_grob(sites[i], color = "red", face = "bold", size = 14)))
}
```

Chronic Overprediction for The Madison Sample is likely due to the more frequent sample collection leading to less change per sample.

```{r}
modelData=MergedData%>%
  group_by(Site)%>%
  filter(!is.na(Model))%>%
  mutate(roll=rollmean(Cases/Tests,k = 7, fill = NA))%>%
  mutate(der=c(diff(roll),NA))


linearMod <- lm(der ~ Model, data=modelData)
linearMod

TempData=AdjustedData%>%
  group_by(Site)%>%
  filter(Date==StartFeces)%>%
  select(Site,startCases=Cases,startTest=Tests)%>%
  merge(modelData,by="Site")

FitedData=TempData%>%
  mutate(fitted =  linearMod[[1]][2]*Model+linearMod[[1]][1],resid = fitted-der,Aresid=abs(resid))%>%
  group_by(Site)%>%
  mutate(orig=cumsum(ifelse(is.na(der), 0, der)) + der*0,origPre=cumsum(ifelse(is.na(fitted), 0, fitted)) + fitted*0)%>%
  mutate(orig=orig+(startCases/startTest),origPre=origPre+(startCases/startTest))%>%
  mutate(residFin=origPre-roll)%>%
  mutate(TotalVarience=roll-mean(roll,na.rm=T))

FitedData%>%
    ggplot()+aes(x=Date)+
    geom_point(aes(y=der,color="change from last week of % positive"))+
    geom_point(aes(y=fitted,color="Model"))+
    ylab("%positive")+
    facet_grid(~Site)+
    scale_x_date(limits = c(StartFeces,EndFeces))


FitedData%>%
    ggplot()+aes(x=der)+
    geom_point(aes(y=fitted),color="red")+
    geom_abline(aes(slope =1,intercept=0))+
    ylab("model predictions")+
    xlab("change from last week of % positive")+
    facet_grid(~Site)
```


R^2 for the model

```{r}
FitedData%>%
  filter(!is.na(resid),!is.na(TotalVarience))%>%
  summarize(R20=1-(sum(resid^2)/sum(TotalVarience^2)))
```

Convert model of change of % positive tests into model of % positive test by summing previous values.

```{r}


FitedData%>%
    ggplot()+aes(x=Date)+
    geom_point(aes(y=origPre,color="Model Intigrated"))+
    geom_point(aes(y=Cases/Tests,color="%Positive"))+
    geom_point(aes(y=orig,color="Reconstructed % Positive"))+
    ylab("%positive")+
    facet_grid(~Site)+
    scale_x_date(limits = c(StartFeces,EndFeces))


FitedData%>%
    ggplot()+aes(x=roll)+
    geom_abline(aes(slope =1,intercept=0))+
    #geom_point(aes(y=orig,color="Reconstructed % Positive"))+
    geom_point(aes(y=origPre),color="red")+
    ylab("Intigrated model predictions")+
    xlab("%positive")+
    facet_grid(~Site)

```


R^2 for the integrated model. Terrible prediction for Madison is due to the chronic overprediction resulting from its more frequent sampling. Other sampling locations have otherwise unexplained inaccuracies

```{r}
FitedData%>%
  filter(!is.na(residFin),!is.na(TotalVarience))%>%
  summarize(R20=1-(sum(residFin^2)/sum(TotalVarience^2)))
```


